From DSL to HPC Component-Based Runtime:\\ A Multi-Stencil DSL Case Study

### 2015-08-02 WOLFHPC15/article/ms_msc.tex
4.1

As mentionned in the previous section, the Multi Stencil Language (MSL) is an agnostic descriptive language for multi-stencil simulations. Six main sections are required in a MSL description, as the six tuples of the formal definition of a multi-stencil program: the mesh description; the mesh entities description; the computation domains description and their dependencies; the data description; the time loop description; and finally, the computations description. Figure~\ref{fig:grammar} shows the grammar of MSL. Lines 1 to 9 define what is a MSL program with the different parts mentionned above. The remaining of this section describes those different parts of a MSL program and presents an example.
We can note that, as MSL is a descriptive language, almost all terminals \texttt{meshid}, \texttt{meshdom}, \texttt{compdomid}, \texttt{dataid}, \texttt{compid} and \texttt{neighborid}, are string identifiers. Only \texttt{num} is an integer terminal.

\subsubsection*{Mesh and Mesh Entities}
As illustrated in the first line of the grammar in Figure~\ref{fig:grammar}, a mesh is directly defined by a string identifier. Then, mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}). In the example given in Figure~\ref{fig:mslex} a Cartesian mesh, called \texttt{cart}, and three mesh entities, called \texttt{cell}, \texttt{edgex} and \texttt{edgey}, are defined.

\subsubsection*{Computation Domains and Their Dependencies}
A computation domain can be seen as a subpart of a mesh entity which will be used during one or more computation. For each computation domain, the mesh entity from which it is built is indicated (lines 3-4 and 12-14 in Figure~\ref{fig:grammar}). By default two computation domains always intersect except when an independency relation is mentionned in the section \emph{independent} of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}). In the example of Figure~\ref{fig:mslex}, on lines 4 to 6, a whole computation domain is defined for each mesh entity, called \texttt{allcell}, \texttt{alledgex} and \texttt{alledgey}, and on lines 7 and 8 two subparts of the mesh entity \texttt{edgex} are also defined, called \texttt{part1edgex} and \texttt{part2edgex}. Finally, line 10 means that \texttt{part1edgex} and \texttt{part2edgex} do not intersect.

\subsubsection*{Data and Time}
A data is a quantity to simulate and is mapped onto a mesh entity, as illustrated in lines 7 and 18-20 of Figure~\ref{fig:grammar}. The \texttt{time} section simply indicates a number of iterations to perform in the simulation. In the example of Figure~\ref{fig:mslex} are defined ten data and 500 iterations (lines 11-22). An extension could be to support a convergence criterion.

\subsubsection*{Computation Description}
Finally, the language follows the two definitions of a stencil kernel and a local computation given in Section~\ref{sect:multistencil}. However, as MSL does not handle numerical expressions, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} is not used in the language. The exact code of the computation as to be written into the final ready-to-fill parallel pattern of the simulation, using external languages or interfaces (\cf Section~\ref{sect:component}). Moreover, to propose an intuitive syntax, information are given in a different order than in formal definitions.
Thus, the description of a computation, as indicated in lines 24 and 25 of Figure~\ref{fig:grammar}, consists in first the identifier of the data to compute ($w$ in the formal description), in brackets the identifier of the computation domain on which the computation is applied ($d$ in the formal definition), an identifier for the computation kernel, and the set of identifiers for the data read by the computation, with parenthesis. For each data read by the computation brackets indicate the stencil shape identifier used on the data ($R$ in the formal definition of a stencil). If the computation is local no brackets appear ($R_l$ in the formal definition). In the example of Figure~\ref{fig:mslex}, nine computations are described. For example, on line 24, the data \texttt{b} is computed on the computation domain \texttt{allcell} by the computation \texttt{c\_0} which read data \texttt{a} without neighborhood shape.

>>>>

As mentioned in the previous section, the Multi Stencil Language (MSL) is an agnostic descriptive language for multi-stencil simulations. Six main sections are required in a MSL description that match the six tuples of the formal definition of a multi-stencil program: \textbf{1.}\,the mesh description, \textbf{2.}\,the mesh entities description, \textbf{3.}\,the computation domains description and their dependencies, \textbf{5.}\,the data description, \textbf{5.}\,the time loop description, and finally \textbf{6.} the computations description.
Figure~\ref{fig:grammar} shows the grammar of MSL. Lines 1 to 9 define what is a MSL program with the different parts mentioned above. The remaining of this section describes those different parts of a MSL program and presents an example.
We can note that, as MSL is a descriptive language, almost all terminals \texttt{meshid}, \texttt{meshdom}, \texttt{compdomid}, \texttt{dataid}, \texttt{compid} and \texttt{neighborid}, are string identifiers. Only \texttt{num} is an integer terminal.

\subsubsection*{Mesh and Mesh Entities}
As illustrated in the first line of the grammar in Figure~\ref{fig:grammar}, a mesh is simply defined by a string identifier. Then, mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}). Figure~\ref{fig:mslex} presents an example defining a Cartesian mesh, called \texttt{cart}, and three mesh entities, called \texttt{cell}, \texttt{edgex} and \texttt{edgey}.

\subsubsection*{Computation Domains and Their Dependencies}
A computation domain can be seen as a subpart of entities of a mesh which will be used during one or more computation. For each computation domain, the mesh entity from which it is built is indicated (lines 3-4 and 12-14 in Figure~\ref{fig:grammar}). By default two computation domains always intersect except when an independence relation is mentioned in the section \emph{independent} of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
In the lines 4 to 6 of Figure~\ref{fig:mslex}, computation domains are defined for each mesh entity, called \texttt{allcell}, \texttt{alledgex} and \texttt{alledgey}; they represent the whole domain.
On lines 7 and 8 two other subparts of the mesh entity \texttt{edgex} are defined, called \texttt{part1edgex} and \texttt{part2edgex}; they represent the boundaries of the domain.
On line 10 the independence relation between \texttt{part1edgex} and \texttt{part2edgex} is specified since they do not intersect.

\subsubsection*{Data and Time}
A data is a quantity to simulate. It is mapped on mesh entities, as illustrated in lines 7 and 18-20 of Figure~\ref{fig:grammar}. The \texttt{time} section simply indicates a number of iterations to perform in the simulation. In the example of Figure~\ref{fig:mslex} ten data identifiers are defined and 500 iterations (lines 11-22). %An extension could be to support a convergence criterion.

\subsubsection*{Computation Description}
The last part of the language contains the specifications of stencil kernels and local computations as defined in Section~\ref{sect:multistencil}.
This description does not contain a direct expression of the numerical expressions.
Instead, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} simply corresponds to an identifier that references an implementation in an external language  (\cf Section~\ref{sect:component}).
The specification starts by the identifier of the data to compute ($w$ in the formal description) with the computation domain between brackets ($d$ in the formal definition).
Then, after the equal sign is specified the kernel identifier.
Finally, between the parenthesis is the list of identifiers of data read by the computation with their stencil shape between brackets ($R$ in the formal definition).
If the computation is local, no brackets appear ($R_l$ in the formal definition). In the example of Figure~\ref{fig:mslex}, nine computations are described. On line 24, the data \texttt{b} is computed on the computation domain \texttt{allcell} by the computation \texttt{c\_0} which read data \texttt{a} without neighborhood shape.
